---
title: "Course Project Phase 1"
author: "Kayleigh Cotton"
date: "2025-06-01"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r Loading Packages and importing dataset}
library(tidyverse)
library(tidymodels)
library(glmnet)
library(GGally)
library(ggcorrplot)
library(MASS)
library(car)
library(lubridate)
library(lmtest)
library(splines)
library(mice) #package for imputation
library(VIM) #visualizing missingness
library(naniar) #visualizing missingness
library(ranger) #for random forests
library(randomForest) #also for random forests
library(caret)
library(skimr) #alternative way to view dataset summaries
library(UpSetR)
train_set <- read_csv("~/Desktop/BAN 502- Predictive Analytics/Course Project/train.csv")
```

```{r Summary of the Data}
str(train_set)
summary(train_set)
skim(train_set)
ggcorr(train_set)
```

Row wise deletion of any row with at least one missing variable
```{r Row Wise Deletion}
train_rowdel = train_set %>% drop_na() 
skim(train_rowdel)
```
This removed 14,387 rows or 54.15% of the rows.


```{r Imputing the missing data}
set.seed(1234)
imp_train = mice(train_set, m=5, method = 'pmm',printFlag = FALSE)
summary(imp_train)
```

```{r}
train_set_complete = complete(imp_train)
str(train_set_complete)
summary(train_set_complete)
skim(train_set_complete)
ggcorr(train_set_complete)
```

### **Phase 1 Project Description**
For Phase 1 you will conduct a thorough exploratory/descriptive analysis of the dataset. Please
**DO NOT** build any predictive models (e.g., logistic regression, trees, etc.) in this phase.
Assume that your "audience" for this work are non-technical decision-makers.
### **Phase 1 Deliverables:**
There are two deliverables for Phase 1.

Deliverable 1: A PowerPoint presentation summarizing your findings from Phase 1. The
presentation should be no more than seven slides (including a title slide). Your findings should
indicate which variables may be strong predictors of the "failure" variable (this is your response
variable) and any other interesting descriptive findings. You should include charts/visuals in the
presentation. There should be NO VISIBLE R CODE in this deliverable. As noted above, you should
assume that the target audience for the deliverable is relatively "non-technical." NOTE: If you create
any variables (i.e., by combining or modifying existing variables), please note this.

Deliverable 2: A knitted Word document of your Phase 1 R work.

