---
title: "Module 3 Quiz 1"
author: "Kayleigh Cotton"
date: "2025-05-24"
output: word_document
editor_options: 
  chunk_output_type: console
---

###**Model Validation**
Libraries: For this assignment you will need the following libraries: tidyverse, lubridate, and tidymodels.
Before beginning the assignment tasks, read-in the “bike_cleaned.csv” file into a data frame called “bike”. This
is the same data that you used in the Module 2 Multiple Linear Regression and Special Issues assignment. As
we did in that assignment you should convert “dteday” from a character variable to a date variable. Convert
the remaining character variables to factors. You can do this one variable at a time or use a “mutate_if”.
Finally, convert the “hr” variable into a factor.

```{r Libraries and Importing}
library(tidyverse)
library(lubridate)
library(tidymodels)
bike <- read_csv("~/Desktop/BAN 502- Predictive Analytics/Module 3/bike_cleaned-4.csv")
bike = bike %>% mutate(dteday = mdy(dteday))
bike = bike %>% mutate_if(is.character, as_factor)
bike = bike %>% mutate(hr = as_factor(hr))
```

```{r Split into Training and Testing Sets} 
set.seed(1234)
bike_split = initial_split(bike, prop = 0.70, strata = count)
train_bike = training(bike_split)
test_bike = testing(bike_split)
```

Question 1: Split the data into training and testing sets. Your training set should have 70% of the data.
Use a random number (set.seed) of 1234. Your split should be stratified by the “count” variable.
How many rows of data are in the training set? I know it’s probably a bit annoying to keep answering this
question about the number of rows, but it’s helpful to be able to validate that your split code is correct before
proceeding :)
**12163**

Question 2 Stratifying the split by the “count” variable serves what purpose?
A. Stratifying by “count” ensures that unusual values for “count” are eliminated
**B. Stratifying by “count” ensures that “count” is similarly represented in both the training and testing sets**
C. Stratifying by “count” ensures that the training set contains the “count” variable
D. None of the above

```{r Linear Regression Model and Performance}
bike_recipe = recipe(count ~ season + mnth + hr + holiday + weekday + temp + weathersit, train_bike)

lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 

lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(bike_recipe)

lm_fit = fit(lm_wflow, train_bike)

summary(lm_fit$fit$fit$fit)

predict_train_bike <- predict(test_bike) %>% bind_cols(test_bike) %>% metrics(truth = count, estimate = .pred)

lm_fit %>% predict(test_bike) %>% bind_cols(test_bike) %>% metrics(truth = count, estimate = .pred)
```

Question 3: Build a linear regression model (using the training set) to predict “count” using the variables
“season”, “mnth”, “hr”, “holiday”, and “weekday”, “temp”, and “weathersit”.
What is the adjusted R-squared value (to four digits) of the resulting model?
**0.6209**

Question 4: Use the predict functions to make predictions (using your model from Question 3) on the
training set. Hint: Be sure to store the predictions in an object, perhaps named “predict_train”
or similar. Develop a histogram of the predictions (Hint: The predictions are likely stored in a variable
called “.pred” in your predictions object).
Select the statements below that are likely true about the distribution of predictions?
**A. The maximum number of rides predicted for an hour is around 600**
B. The average number of rides predicted per hour is around 450
**C. Some predictions for the number of rides in an hour are negative**
D. None of these statements are true

Question 5: Determine the performance of your model on the testing set.
What is the R-squared value (to four decimal places) of your model on the testing set? REMINDER: DO
NOT build a model on the testing set. Use your model that was developed on the training set.
**0.6270**